### 2018-01-10 RFCs/HDF5/VOL/user_guide/external.tex
updated LateX and general editing of document
Former-commit-id: 2e4326ee524db559b5b9ab32ee07a03caa786f9a

\section{Creating and Using an External Plugin}
External plugins are developed outside of the HDF5 library and do not
use any internal HDF5 private functions. They do not require to be
shipped with the HDF5 library, but can just link to it from userspace
just like an HDF5 application. If an external VOL plugin is installed on system as a shared library or a DLL, the library can search for it and load it dynamically to be used by an application.

\subsection{Creating an External Plugin}
Developing an external plugin is similar to developing an internal plugin. Refer to Section~ref{sec:vol} for details on plugin creation. The important thing to keep in mind is that external plugins cannot use internal HDF5 features. The ``value'' field for an external plugin should be a positive integer greater than 128. The name should be unique across all plugins registered with the library.

In order for the HDF5 library to be able to load an external plugin dynamically, the plugin developer has to define two public routines with the following name and signature:
\begin{lstlisting}
H5PL_type_t H5PLget_plugin_type(void)
const void *H5PLget_plugin_info(void)
\end{lstlisting}
{\tt H5PLget\_plugin\_type} should return the library type which should always be {\tt H5PL\_TYPE\_VOL}. {\tt H5PLget\_plugin\_info} should return a pointer to the plugin structure defining the VOL plugin with all the callbacks. For example, consider an external plugin defined as:
\begin{lstlisting}
static const H5VL_class_t H5VL_log_g = {
    1,	/* version */
    502,	/* value */
    "log",	/* name */
    ...
}
\end{lstlisting}
The plugin would implement the two routines as:
\begin{lstlisting}
H5PL_type_t H5PLget_plugin_type(void) {return H5PL_TYPE_VOL;}
const void *H5PLget_plugin_info(void) {return &H5VL_log_g;}
\end{lstlisting}

@@

\subsection{Using an External Plugin}
Unlike internal plugins, external plugins cannot create an API routine for applications to use to set the VOL plugin in the file access property list. External plugin developers can however provide a wrapper routine that registers the plugin with {\tt H5VLregister()} instead of having the application call that directly with the pointer to the plugin class structure. Both ways are possible if the external plugin used is in the application's user space. External plugins provided by third pary vendors that are not in the application's user space but are installed on the system with a shared library or DLL can be registered by the application using {\tt H5VLregister\_name()} provided that the application knows the plugin's name in advance.

The registration operation will return a global identifier for the registered plugin. Applications can query the library for that identifier using the plugin name with {\tt H5VLget\_plugin\_id()}. This identifier is used to initialize the plugin if it requires an initialization phase. A third party library could use that ID to create a plugin that stacks or mirrors on top of an internal plugin using that identifier (refer to section~\ref{sec:stack} for more details on stacking and mirroring).

@@

Finally the application is required to un-register the plugin from the library when access to the container(s) is terminated using {\tt H5VLunregister}.

>>>>

\section{Creating and Using an External Plugin}
External plugins are developed outside of the HDF5 library and do not
use any internal HDF5 private functions. They do not require to be
shipped with the HDF5 library, but can just link to it from userspace
just like an HDF5 application. If an external VOL plugin is installed on the system as a shared library or a DLL, the library can search for it and load it dynamically to be used by an application.

\subsection{Creating an External Plugin}
Developing an external plugin is similar to developing an internal plugin. Refer to Section~\ref{sec:vol} for details on plugin creation. The important thing to keep in mind is that external plugins cannot use internal HDF5 features. The ``value'' field for an external plugin should be a positive integer greater than 128. The name should be unique across all plugins registered with the library.

For the HDF5 library to be able to load an external plugin dynamically, the plugin developer has to define two public routines with the following name and signature:
\begin{lstlisting}
H5PL_type_t H5PLget_plugin_type(void)
const void *H5PLget_plugin_info(void)
\end{lstlisting}
{\tt H5PLget\_plugin\_type} should return the library type which should always be {\tt H5PL\_TYPE\_VOL}. {\tt H5PLget\_plugin\_info} should return a pointer to the plugin structure defining the VOL plugin with all the callbacks. For example, consider an external plugin defined as:
\begin{lstlisting}
static const H5VL_class_t H5VL_log_g = {
    1,	/* version */
    502,	/* value */
    "log",	/* name */
    ...
}
\end{lstlisting}
The plugin would implement the two routines as:
\begin{lstlisting}
H5PL_type_t H5PLget_plugin_type(void) {return H5PL_TYPE_VOL;}
const void *H5PLget_plugin_info(void) {return &H5VL_log_g;}
\end{lstlisting}

@@

\subsection{Using an External Plugin}
Unlike internal plugins, external plugins cannot create an API routine for applications to use to set the VOL plugin in the file access property list. External plugin developers can, however, provide a wrapper routine that registers the plugin with {\tt H5VLregister()} instead of having the application call that directly with the pointer to the plugin class structure. Both ways are possible if the external plugin used is in the application's userspace. External plugins provided by third party vendors that are not in the application's userspace but are installed on the system with a shared library or DLL can be registered by the application using {\tt H5VLregister\_name()} provided that the application knows the plugin's name in advance.

The registration operation will return a global identifier for the registered plugin. Applications can query the library for that identifier using the plugin name with {\tt H5VLget\_plugin\_id()}. This identifier is used to initialize the plugin if it requires an initialization phase. A third party library could use that ID to create a plugin that stacks or mirrors on top of an internal plugin using that identifier (refer to Section~\ref{sec:stack} for more details on stacking and mirroring).

@@

Finally, the application is required to un-register the plugin from the library when access to the container(s) is terminated using {\tt H5VLunregister}.

### 2018-01-10 RFCs/HDF5/VOL/user_guide/new_api.tex
updated LateX and general editing of document
Former-commit-id: 2e4326ee524db559b5b9ab32ee07a03caa786f9a

\begin{lstlisting}
hid_t H5VLregister(const H5VL_class_t *cls);
\end{lstlisting}
Registers a user defined VOL plugin with the HDF5 library and returns an identifier for that plugin to be used if successful, otherwise returns a negative value. This function is used when the application has direct access to the plugin it wants to use and is able to obtain a pointer for the plugin structure to pass to the HDF5 library. The plugin must be unregistered before the application exists with {\tt H5VLunregister()}.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt cls} & (IN): A pointer to the plugin structure to register.\\
\end{tabular}

\begin{lstlisting}
hid_t H5VLregister_by_name(const char *plugin_name);
\end{lstlisting}
Registers a VOL plugin with the HDF5 library given the name of the plugin and returns an identifier for that plugin to be used if successful, otherwise returns a negative value. If the plugin is already registered the library will create an identifier for it and returns it to the user; otherwise the library will look for available third party plugins and dynamically load them and registers them for use, then returns an identifier for the plugin. Third party plugins must be installed on the system as a shared library or DLL. The plugin must be unregistered before the application exists with {\tt H5VLunregister()}.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to search for an register.\\
\end{tabular}

\begin{lstlisting}
herr_t H5VLunregister(hid_t plugin_id);
\end{lstlisting}
Unregisters a plugin from the library and return a positive value on success otherwise return a negative value.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt plugin\_id} & (IN): A valid identifier of the plugin to unregister.\\
\end{tabular}

\begin{lstlisting}
htri_t H5VLis_registered(const char *name);
\end{lstlisting}
Checks if a VOL plugin is registered with the library given the plugin name and returns TRUE/FALSE on success, otherwise it returns a negative value.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to check for.\\
\end{tabular}

\begin{lstlisting}
ssize_t H5VLget_plugin_name(hid_t obj_id, char *name/*out*/, size_t size);
\end{lstlisting}
Retrieves the name of a VOL plugin given and object identifier that was created/opened with it. On success, the name length is returned.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt obj\_id} & (IN): The object identifier to check.\\
  {\tt name} & (OUT): Buffer pointer to put the plugin name. If NULL, the library just returns the size required to store the plugin name.\\
  {\tt size} & (IN): the size of the passed in buffer.\\
\end{tabular}

\begin{lstlisting}
hid_t H5VLget_plugin_id(const char *name);
\end{lstlisting}
Given a plugin name that is registered with the library, this function returns an identifier for the plugin. If the plugin is not registered with the library, a negative value is returned. The identifier must be released with a call to {\tt H5VLclose()}.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to check for.\\
\end{tabular}

\begin{lstlisting}
herr_t H5VLclose(hid_t plugin_id);
\end{lstlisting}
Shuts down access to the plugin that the identifier points to and release resources associated with it.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): A valid identifier of the plugin to close.\\
\end{tabular}

\begin{lstlisting}
  hid_t H5VLobject_register(void *obj, H5I_type_t obj_type, hid_t plugin_id);
\end{lstlisting}
Creates an HDF5 identifier for an object given its type and an identifier of the plugin that was used in creating it. This function is typically used by external plugins wanting to wrap an {\tt hid\_t} around a plugin specific object to return to a user callback for functions like {\tt H5Literate, H5Lvisit, H5Ovisit, H5Aiterate} etc... A Negative value is returned on failure.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt obj} & (IN): pointer to the object to register.\\
 {\tt obj\_type} & (IN): The type of the object ({\tt H5I\_FILE, H5I\_DATASET, H5I\_GROUP, H5I\_ATTRIBUTE, H5I\_DATATYPE}).\\
 {\tt plugin\_id} & (IN): identifier of the vol plugin the object is associated with.\\
\end{tabular}

\begin{lstlisting}
  herr_t H5VLget_object(hid_t obj_id, void **obj); 
\end{lstlisting}
Retrieves a pointer to the VOL object from an HDF5 file or object identifier.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt obj\_id} & (IN): identifier of the object to dereference.\\
 {\tt obj} & (OUT): buffer to store the pointer of the VOL object associated with the object identifier.\\
\end{tabular}

\begin{lstlisting}
herr_t H5VLinitialize(hid_t plugin_id, hid_t vipl_id);
\end{lstlisting}
Initializes access to the VOL plugin represented by {\tt plugin\_id} with initialization parameters set as properties in the the property list {\tt vipl\_id}. Some plugins may not require initialization. For more information about plugin initialization and what properties are required to be set in the property list refer to the actual plugin documentation.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): identifier of VOL plugin to initialize.\\
 {\tt vipl\_id} & (IN): VOL initialize property list containing properties required for plugin initialization.\\
\end{tabular}

\begin{lstlisting}
herr_t H5VLterminate(hid_t plugin_id, hid_t vtpl_id);
\end{lstlisting}

Terminates access to the VOL plugin represented by {\tt plugin\_id} with termination parameters set as properties in the the property list {\tt vtpl\_id}. Some plugins may not require termination. For more information about plugin termination and what properties are required to be set in the property list refer to the actual plugin documentation.

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): identifier of VOL plugin to terminate.\\
 {\tt vtpl\_id} & (IN): VOL terminate property list containing properties required for plugin termination.\\
\end{tabular}

>>>>

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
hid_t H5VLregister(const H5VL_class_t *cls);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt cls} & (IN): A pointer to the plugin structure to register.\\
\end{tabular}
\end{mdframed}

Registers a user defined VOL plugin with the HDF5 library and returns an identifier for that plugin to be used if successful, otherwise returns a negative value. This function is used when the application has direct access to the plugin it wants to use and is able to obtain a pointer for the plugin structure to pass to the HDF5 library. The plugin must be unregistered before the application exists with {\texttt H5VLunregister()}.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
hid_t H5VLregister_by_name(const char *plugin_name);
\end{lstlisting}
\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to search for an register.\\
\end{tabular}
\end{mdframed}
Registers a VOL plugin with the HDF5 library given the name of the plugin and returns an identifier for that plugin to be used if successful, otherwise returns a negative value. If the plugin is already registered the library will create an identifier for it and returns it to the user; otherwise the library will look for available third party plugins and dynamically load them and registers them for use, then returns an identifier for the plugin. Third party plugins must be installed on the system as a shared library or DLL. The plugin must be unregistered before the application exists with {\tt H5VLunregister()}.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
herr_t H5VLunregister(hid_t plugin_id);
\end{lstlisting}
\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt plugin\_id} & (IN): A valid identifier of the plugin to unregister.\\
\end{tabular}
\end{mdframed}
Unregisters a plugin from the library and return a positive value on success otherwise return a negative value.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
htri_t H5VLis_registered(const char *name);
\end{lstlisting}
\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to check for.\\
\end{tabular}
\end{mdframed}
Checks if a VOL plugin is registered with the library given the plugin name and returns TRUE/FALSE on success, otherwise it returns a negative value.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
ssize_t H5VLget_plugin_name(hid_t obj_id, char *name/*out*/, size_t size);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt obj\_id} & (IN): The object identifier to check.\\
  {\tt name} & (OUT): Buffer pointer to put the plugin name. If NULL, the library just returns the size required to store the plugin name.\\
  {\tt size} & (IN): the size of the passed in buffer.\\
\end{tabular}
\end{mdframed}
Retrieves the name of a VOL plugin given and object identifier that was created/opened with it. On success, the name length is returned.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
hid_t H5VLget_plugin_id(const char *name);
\end{lstlisting}
\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
  {\tt name} & (IN): The plugin name to check for.\\
\end{tabular}
\end{mdframed}
Given a plugin name that is registered with the library, this function returns an identifier for the plugin. If the plugin is not registered with the library, a negative value is returned. The identifier must be released with a call to {\tt H5VLclose()}.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
herr_t H5VLclose(hid_t plugin_id);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): A valid identifier of the plugin to close.\\
\end{tabular}
\end{mdframed}
Shuts down access to the plugin that the identifier points to and release resources associated with it.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
  hid_t H5VLobject_register(void *obj, H5I_type_t obj_type, hid_t plugin_id);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt obj} & (IN): pointer to the object to register.\\
 {\tt obj\_type} & (IN): The type of the object ({\tt H5I\_FILE, H5I\_DATASET, H5I\_GROUP, H5I\_ATTRIBUTE, H5I\_DATATYPE}).\\
 {\tt plugin\_id} & (IN): identifier of the vol plugin the object is associated with.\\
\end{tabular}
\end{mdframed}
Creates an HDF5 identifier for an object given its type and an identifier of the plugin that was used in creating it. This function is typically used by external plugins wanting to wrap an {\tt hid\_t} around a plugin specific object to return to a user callback for functions like {\tt H5Literate, H5Lvisit, H5Ovisit, H5Aiterate} etc... A Negative value is returned on failure.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
  herr_t H5VLget_object(hid_t obj_id, void **obj); 
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt obj\_id} & (IN): identifier of the object to dereference.\\
 {\tt obj} & (OUT): buffer to store the pointer of the VOL object associated with the object identifier.\\
\end{tabular}
\end{mdframed}
Retrieves a pointer to the VOL object from an HDF5 file or object identifier.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
herr_t H5VLinitialize(hid_t plugin_id, hid_t vipl_id);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): identifier of VOL plugin to initialize.\\
 {\tt vipl\_id} & (IN): VOL initialize property list containing properties required for plugin initialization.\\
\end{tabular}
\end{mdframed}
Initializes access to the VOL plugin represented by {\tt plugin\_id} with initialization parameters set as properties in the the property list {\tt vipl\_id}. Some plugins may not require initialization. For more information about plugin initialization and what properties are required to be set in the property list refer to the actual plugin documentation.\bigskip

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
herr_t H5VLterminate(hid_t plugin_id, hid_t vtpl_id);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{10cm}}
 {\tt plugin\_id} & (IN): identifier of VOL plugin to terminate.\\
 {\tt vtpl\_id} & (IN): VOL terminate property list containing properties required for plugin termination.\\
\end{tabular}
\end{mdframed}
Terminates access to the VOL plugin represented by {\tt plugin\_id} with termination parameters set as properties in the the property list {\tt vtpl\_id}. Some plugins may not require termination. For more information about plugin termination and what properties are required to be set in the property list refer to the actual plugin documentation.
